Zaca Fire Type Conversion

### 2020-10-05 Model Data Sheets and Code/Manuscript_Code.Rmd
final code (messy but will clean up later...hopefully)

library(tidyverse)
library(raster)
library(randomForest)
#library(VSURF)
library(ALEPlot)
library(tuneRanger)

@@

### Extract Values

@@

### Stepwise Variable Selection

@@

#Extract prediction and response variables
mcf.predictors <- mixed_conifer_env[,2:ncol(mixed_conifer_env)]
mcf.response <- mixed_conifer_env[,1]

@@

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.pred])

@@

# Tuning Parameters

@@

replace: dataframe and response variable

tunRanger can also be used for ntree, but studies have shown that after a certain number of trees, usually a couple hundred, there is no benefit on accuracy/model fit of increasing ntree, so i fgiure just using a large number will get the same result as whatever number tuneRangere spits out. 
```{r}
set.seed(1127)

# run tuneRanger
mcf_task <- makeClassifTask(data = dataframe, target = "y variable")
tune_output <- tuneRanger(mcf_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

@@

### Random Forest

#Create random forest using selected variables
mixed_conifer_rf <- randomForest(RdNBR_2007 ~ var1 + var2 + var3, 
                                 data = mixed_conifer_env, 
                                 importance = TRUE, 
                                 na.action = na.omit)

@@

>>>>

library(tidyverse)
library(raster)
library(randomForest)
library(VSURF)
library(ALEPlot)
library(tuneRanger)
library(iml)
library(ggpubr)

@@

###... Extract Values

@@

###... Stepwise Variable Selection

@@

#Extract prediction and response variables
mcf.predictors <- mcf_model[,2:ncol(mcf_model)]
mcf.response <- mcf_model[,1]

@@

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.thres])
names(mcf.predictors[mcf.vsurf$varselect.interp])
names(mcf.predictors[mcf.vsurf$varselect.pred])

```
prediction step, variables selected:all

interp step: variables se;ected: "vpd_aug"         "slope"           "rain"            "tmax"            "tmin_jan"        "sr" "aug_tmax2006"    "tmin_jan2007"    "precip_2006"     "aug_max_vpd2006" "ndmi_2006"       "evi_2006"  "tslf"

prediction step, variables selected: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "sr"           "aug_tmax2006"  "tmin_jan2007" "ndmi_2006"    "evi_2006"     "tslf"

so climate variables are performing better than the weather variables (maybe bc the climate data is avail in higher resolution so there is more variability in values?)

removed aug_max_vpd2006 and precip_2006 and reran VSURF

interp step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "tmin_jan2007"

predicition step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan2007"

###... Load data

the above code wouldnt work, so I got evi and min jan temp in the aca mcf mortality model results.rmd and loaded the old files ive been using
```{r}
mcf <- read.csv(here::here("Model Data Sheets and Code", 'Zaca MCF variables for model.csv')) %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(two_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
  mutate(two_class_mortality = as.factor(two_class_mortality),
         two_class_mortality= fct_relevel(two_class_mortality, levels=c("NLM", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 640 ~ "NLM",
    rdnbr_2007 < 850 ~ "High",
    rdnbr_2007 >= 850 ~ "Type Converted")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, levels=c("NLM","High", "Type Converted"))) %>% 
  #   mutate(rdnbr_qualitative2 = case_when(
  #   rdnbr_qualitative %in% "High" ~ "High",
  #   rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
  # mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2),
  #        rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, levels=c("NLM", "High")))%>% 
  mutate(rdnbr_tc = case_when(
    rdnbr_2007 < 850 ~ "Not_TC",
    rdnbr_2007 >= 850 ~ "TC"
  )) %>% 
  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, levels=c("Not_TC", "TC")))

mcf_tc_samplesize <- mcf %>% group_by(tc_mortality) %>%count() #type converted n = 99
mcf_tc_samplesie <- mcf %>% group_by(rdnbr_tc) %>%dplyr::count() # type converted n = 55 (almost too small to use)
mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::count() # high=66, low= 585, moderate=272, type converted n = 84 
mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::summarise(acres = sum(Shape_Area/4047)) %>%dplyr::summarise(tot_acres = sum(acres))

ggplot(mcf, aes(x = rdnbr_tc, y = rdnbr_2007)) +geom_jitter(aes(color=tc_mortality))
ggplot(mcf, aes(x=dnbr_2007, y=rdnbr_2007)) +geom_point(aes(color=tc_mortality)) + geom_hline(yintercept=875)
ggplot(mcf, aes(ndmi_2006, y=evi_2006)) +geom_point()
ggplot(mcf, aes(dem_10m_zaca, y=tmin_jan2007)) +geom_point()
mod <- lm(ndmi_2006~evi_2006, data=mcf) # R2=0.18
mod1 <- lm(dem_10m_zaca~tmin_jan2007, data=mcf) # R2=0.04
mod2 <- lm(dem_10m_zaca~cwd, data=mcf) # R2=0.47
mod3 <- lm(dem_10m_zaca~rain, data=mcf) # R2=0.498
mod4 <- lm(dem_10m_zaca~tmax, data=mcf) # R2=0.5
mod5 <- lm(dem_10m_zaca~vpd_aug, data=mcf) # R2=0.56
mod6 <- lm(tmax~vpd_aug, data=mcf) # R2=0.95
mod7 <- lm(cwd~vpd_aug, data=mcf) # R2=0.53
mod8 <- lm(tmax~cwd, data=mcf) # R2=0.49

mcf %>% dplyr::summarise(min = min(tslf),
                         mean = mean(tslf),
                         median = median(tslf),
                         max = max(tslf), 
                         min_elev = min(dem_10m_zaca),
                         mean_elev = mean(dem_10m_zaca),
                         max_elev = max(dem_10m_zaca))

@@

###... Tuning Parameters

@@

tunRanger can also be used to tune for ntree, but studies have shown that after a certain number of trees, usually a couple hundred, there is no benefit on accuracy/model fit of increasing ntree, so i fgiure just using a large number will get the same result as whatever number tuneRangere spits out. 
```{r}
set.seed(1127)

# select variables to be used in the model
mcf_model <- mcf %>% 
  dplyr::select(rdnbr_tc, tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing

# run tuneRanger
mcf_task <- makeClassifTask(data = mcf_model, target = "rdnbr_tc")
tune_output <- tuneRanger(mcf_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

# 4 rdnbr classes
mcf_model3 <- mcf %>% 
  dplyr::select(rdnbr_qualitative, tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing

# run tuneRanger
mcf_task <- makeClassifTask(data = mcf_model3, target = "rdnbr_qualitative")
tune_output <- tuneRanger(mcf_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

@@

Trials: (2 classes)
- rdnbr cut off: 850; variables included: tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TPI_five, TWI, ndmi_2006, evi_2006, fri; optimal parameters: mtry=6, min.node.size=4; acc = 82%, not tc = 75%, tc=62%, auc=84%; stable VI: kinda
Using this one- rdnbr cut off: 850; variables included: removed tpi, mtry=3,min.node.size=7
- rdnbr cut off: 850; variables included: removed tmin; optimal parameters: mtry=5, min.node.size=10; stable VI: kinda
- rdnbr cut off: 850; tmax, vpd_aug, rain, sr, slope, aspect_transformed, evi_2006, tslf, tmin_jan2007, ndmi_2006; mtry=3,min.node.size=7; stable vi: sure!
- rdnbr cut off: 850; variables included: removed cwd; optimal parameters: mtry=4, min.node.size=2; stable VI: NO! horrible stability. 
- rdnbr cut off: 850; variables included: removed evi_2006; optimal parameters: mtry=4, min.node.size=3; stable VI:kinda
- rdnbr cut off: 850; variables included: removed ndmi_2006; optimal parameters: mtry=6, min.node.size=27; stable VI:kinda

- rdnbr cut off: 875; variables included: all; optimal parameters: mtry=6, min.node.size=15; stable VI:eh, kinda
- rdnbr cut off: 875; variables inlcuded: rdnbr_tc, tmax, vpd_aug, rain, tmin_jan, sr, slope, aspect_transformed, TPI_five, TWI, evi_2006, tslf, cwd, tmin_jan2007, ndmi_2006, precip_2006, aug_max_vpd2006, aug_tmax2006; mtry=7,min.node.size=8; stable VI: NO! horrible stability. 
- rdnbr cut off: 875; variables inlcuded: removed precip_2006 and aug_max_vpd2006 bc of VSURF results; mtry=6,min.node.size=5

- rdnbr cut off: 875; variables inlcuded:tmax, vpd_aug, rain, sr, slope, aspect_transformed, TPI_five, TWI, evi_2006, tslf, cwd, tmin_jan2007, ndmi_2006; mtry=4,min.node.size=3; stable VI: NO! horrible stability.

4 classes
mtry=12,min.node.size=5; stable VI!!; oob not great, TC has lots of confusion with moderate class

3 classes
mtry=9,min.node.size=3; VI is mostly stable!
###... Random Forest
```{r}
for (i in 1:10) {

rfm <- randomForest(rdnbr_qualitative ~., 
                    data = mcf_model3, 
                    #sampsize = c('Not_TC' = 84, 'TC' = 84), 
                    sampsize = c("NLM" = 84,  "High" = 66, "Type Converted" = 84),
                    ntree=1500, 
                    importance=T,
                    replace = T, 
                    mtry = 9,
                    min.node.size = 3)
# OOB and importance
rfm
#importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

```
###... ROC/AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf_model$rdnbr_tc, rfm$votes[,2])
plot(rfm_roc)
auc(rfm_roc)
```
###... ALE
```{r}

mcf_ale <- mcf_model3[-which(names(mcf_model3) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="NLM")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7
#ndmi
nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm1, high1, tc1, nrow=1)
##ggsave(here::here("Graph Images", "mcf ndmi.jpg"))
#theme (panel.background = element_rect(width = 5))

#solar radiation
nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf sr.jpg"))

#precip
nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf precip.jpg"))

#evi
nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf evi.jpg"))

#tmax
nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tmax.jpg"))

#tmin jan 2007
nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))

#vpd aug
nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))

#cwd
nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf cwd.jpg"))

#slope
nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf slope.jpg"))

#tslf
nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tslf.jpg"))

#twi
nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf twi.jpg"))

#aspect
nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf Aspect.jpg"))

ggarrange(nlm1, nlm2,nlm3,nlm4,nlm5,nlm6,nlm7,nlm8,nlm9,nlm10,nlm11,nlm12, nrow = 3, ncol=4)
ggsave(here::here("Graph Images", "mcf all ulm.jpg"), width=8, height = 8)

ggarrange(high1,high2,high3,high4,high5,high6,high7,high8,high9,high10,high11,high12, nrow=3, ncol=4)
ggsave(here::here("Graph Images", "mcf all high.jpg"), width=8, height = 8)

ggarrange(tc1,tc2,tc3,tc4,tc5,tc6,tc7,tc8,tc9,tc10,tc11,tc12, nrow=3, ncol=4)
ggsave(here::here("Graph Images", "mcf all tc.jpg"), width=8, height = 8)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(tc_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'sr', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'tslf', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("mcf tc.jpg", height = 9, width = 9)

@@

###... ALE
```{r}

mcf_ale <- mcf_model[-which(names(mcf_model) == "rdnbr_tc")]

# predictors 
noTC_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_tc, type = "prob", class="Not_TC")
TC_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_tc, type = "prob", class="TC")

#ndvi
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "vpd_aug")) + ggtitle("Not Type Converted") + labs(x="NDVI")
high <- plot(FeatureEffect$new(TC_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="NDVI")
ggarrange(Low, high)

#temp
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "rain")) + ggtitle("Not Type Converted")+ labs(x="Precipitation")
high <- plot(FeatureEffect$new(TC_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="Precipitation")
ggarrange(Low, high)

#precip
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "evi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Outside")
high <- plot(FeatureEffect$new(TC_predictor, feature = "evi_2006")) + ggtitle("Type Converted")+ labs(x="Vegetation Outside")
ggarrange(Low, high)

#precip
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "ndmi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Inside")
high <- plot(FeatureEffect$new(TC_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="Vegetation Inside")
ggarrange(Low, high)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("ale4.jpg", height = 9, width = 9)
```
interesting graphs for evi and ndmi: high evi (which represents denser veg) is associated with low mortality, which is counterintuative, but higher ndmi (which represents moist veg??) is also associated with low mortality.

## Coulter Pines

#### Load data

```{r}
pico <- read_csv("coulter_env_modified.csv") %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 640 ~ "NLM",
    rdnbr_2007 < 850 ~ "High",
    rdnbr_2007 >= 850 ~ "Type Converted")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, levels=c("NLM","High", "Type Converted"))) %>% 
  #filter(Mortality != "bigcone") %>% # 60 stands actually have big cones not pico
  dplyr::select(-Mortality, -type_converted, -ndvi_2006) %>% 
  na.omit(pico) %>%  # 56 stands have NA values (probs all very small stands)
   mutate(rdnbr_qualitative2 = case_when(
    rdnbr_qualitative %in% "High" ~ "High",
    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, levels=c("NLM", "High"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect_30m)/360))/2) %>% 
  mutate(rdnbr_tc = case_when(
    rdnbr_2007 < 850 ~ "Not_TC",
    rdnbr_2007 >= 850 ~ "TC"
  )) %>% 
  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, levels=c("Not_TC", "TC"))) %>% 
  rename_all(funs(str_replace(., "_30m", "")))

pico_tc_samplesize <- pico %>% group_by(rdnbr_tc) %>% dplyr::count() # 
pico_tc_samplesize <- pico %>% group_by(rdnbr_qualitative) %>% dplyr::count() # nlm = 324, high=120, tc=595
pico_tc_samplesize <- pico %>% group_by(rdnbr_qualitative) %>% dplyr::summarise(acres = sum(acres)) %>% dplyr::summarise(total_acres = sum(acres))
mod1 <- lm(dem~tmin_jan2007, data=pico) # R2=0.368
mod2 <- lm(dem~cwd, data=pico) # R2=0.68
mod3 <- lm(dem~rain, data=pico) # R2=0.475
mod4 <- lm(dem~tmax, data=pico) # R2=0.86
mod5 <- lm(dem~vpd_aug, data=pico) # R2=0.455

pico_elev <- pico %>% dplyr::summarise(elev_mean = mean(dem),
                                       min_elev = min(dem),
                                       max_elev =max(dem))

###... Tuning Parameters

```{r}
set.seed(1127)

# select variables to be used in the model
pico_model2 <- pico %>% dplyr::select(rdnbr_tc, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006)

# run tuneRanger
pico_task <- makeClassifTask(data = pico_model2, target = "rdnbr_tc")
tune_output <- tuneRanger(pico_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

## other 
# select variables to be used in the model
pico_model3 <- pico %>% dplyr::select(rdnbr_qualitative, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007)

# run tuneRanger
pico_task <- makeClassifTask(data = pico_model3, target = "rdnbr_qualitative")
tune_output <- tuneRanger(pico_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

rdnbr = 850: tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006; mtry=3,min.node.size=2

3 class: mtry=1,min.node.size=2; VI stable!

###... Random Forest
```{r}

for (i in 1:10) {

rfm <- randomForest(rdnbr_qualitative ~., 
                    data = pico_model3, 
                    #sampsize = c('Not_TC' = 444, 'TC' = 444), 
                    sampsize = c("NLM" = 120,  "High" = 120, "Type Converted" = 120),
                    ntree=1500, 
                    importance=T,
                    replace = T, 
                    mtry = 1,
                    min.node.size = 2)
# OOB and importance
rfm
#importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

#ggplot(pico_model, aes(x=tmin_jan, y=tmin_jan2007)) + geom_point()
#ggplot(pico_model, aes(x=tmax, y=aug_tmax2006)) + geom_point()

###... ALE
```{r}
pico_ale <- pico_model3[-which(names(pico_model3) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="NLM")
high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7

#ndmi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico ndmi.jpg"))
#theme (panel.background = element_rect(width = 5))
